Laplacian envelope

### 2015-06-22 Notes/laplace.tex
reorganised

\begin{figure}[tbp]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv07.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv13.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv21.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv27.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv36.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv51.eps}}
  \end{myenuma}
  \caption[Optimal models]{The models that constitute the numeric envelope in \autoref{fig:envnum}.
  The plot to the left of each frame indicates the memory curve of the model shown to the right in blue, with the numeric envelope for models with the serial topology in red and the proven envelope, \eqref{eq:env}, in green.
  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
  See the video for more details.}\label{fig:envvid}
\end{figure}

@@

Looking at them models that were found by numerical maximisation in \autoref{fig:envnum} (see the video for the models), suggests we look at the particular set of models, shown in \autoref{fig:heuristicmodels}.

At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as $\tau\to\infty$.
At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
Once the length has been reduced to $M=2$, the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.

@@

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0\text{ or }1$:
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap.
This is related to the fact that the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.

>>>>

\begin{figure}[tbp]
  \centering
  % Requires \usepackage{graphicx}
  \begin{myenuma}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv07.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv13.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv21.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv27.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv36.eps}}
  \item\aligntop{\includegraphics[width=0.45\linewidth]{Lenv51.eps}}
  \end{myenuma}
  \caption[Optimal models]{The models that constitute the numeric envelope in \autoref{fig:envnum}.
  The plot to the left of each frame indicates the memory curve of the model shown to the right in blue, with the numeric envelope for models with the serial topology in red and the proven envelope, \eqref{eq:env}, in green.
  The bar graph in the upper right of each frame indicates the transition probabilities from the state to the left of each bar to the state on its right, under potentiation.
  The bar graph in the lower right of each frame indicates the transition probabilities from the state to the right of each bar to the state on its left, under depression.
  The heat map in the centre right of each frame indicates the equilibrium probability distribution for each state.
  See the video for the full set of models.}\label{fig:envvid}
\end{figure}

@@

Looking at them models that were found by numerical maximisation in \autoref{fig:envvid} (see the video for the full set of models), suggests we look at the particular set of models, shown in \autoref{fig:heuristicmodels}.

At one particular time, the model is (\ref{fig:serial_uniform}) the uniform serial model with $M$ states.
At later times, the model has (\ref{fig:serial_sticky}) ``sticky'' end states with low exit transition probabilities, tending to the area maximising model as $\tau\to\infty$.
At earlier times, the model is (\ref{fig:serial_shorten}) progressively shortened by reducing the transition probability to the end states.
Once this reaches zero, we discard the old end state and shorten further by reducing the transition probability to the new end state (as one can see from \autoref{fig:envvid} and the video, this is not quite true, but it is close enough to true to get a reasonable approximation to the envelope).
Once the length has been reduced to $M=2$, the model is not shortened any further and the two-state model with deterministic transitions maximises the memory curve for all earlier times.

@@

We can maximise this \wrt $\varepsilon$.
This results in a complicated quadratic equation for $\varepsilon$ with an even worse solution that we will not reproduce here.
We can find the boundary of the region of applicability by seeing at what value of $s$ is the optimal value of $\varepsilon=0\text{ or }1$:
%
\begin{equation}\label{eq:shortenvalid}
\begin{aligned}
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=1} \propto &\,
    M(\alpha^2-\alpha+1)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^{\frac{M}{2}}-1)^2(\alpha^{M}+1)
      = 0,\\
  \left.\pdiff{\hat{A}(s)}{\varepsilon}\right|_{\varepsilon=0} \propto &\,
    (M-2)(\alpha-1)\alpha^{\frac{M-2}{2}}(\alpha^M-1) - (\alpha^2-\alpha+1)(\alpha^{\frac{M-2}{2}}-1)^2(\alpha^{M-2}+1)
      = 0,\\
    s = &\, \frac{(\alpha-1)^2}{2\alpha},
\end{aligned}
\end{equation}
%
with the $\varepsilon=1$ case providing the lower boundary and $\varepsilon=0$ providing the upper boundary.
If these were really the models that maximised $A(s)$, the region of validity for $M$ would not overlap with that for $M-2$.
In this case, however, there is an overlap.
This is related to the fact that \autoref{fig:envvid} and the video shows that the optimal models do not6 shorten by one state at a time, but several outgoing transition probabilities are reduced at once.
